---
title: "ADA_finalproject"
output: html_document
---
#Install and load all packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages(pacman)
install.packages("table1")
pacman::p_load(tidyverse, ROCR, odds.n.ends, blorr, lmtest, car)
library(table1)
library(foreign)
library(haven)
library(readr)
library(knitr)
library(psych)
library(tidyverse)
library(ggplot2)
library(plyr)
library(dplyr)
library(naniar)
```


#To open dataset
```{r}
CHIS_spss <-
  read.spss(
    "C:\\Users\\adash\\OneDrive\\Documents\\ADA_finalproj_dataset.sav", to.data.frame=TRUE
    )
```

#Rename variable names 
```{r}
names(CHIS_spss)[1:6]<-c("asthma", "smoking", "race", "bmi", "age", "sex")
```

#checking class of outcome variable
```{r}

class(CHIS_spss$asthma)
table(CHIS_spss$asthma)
```

#Recoding asthma variable to 0  & 1 for logistic regression
```{r}
CHIS <- CHIS_spss %>%
  mutate(asthma_binary = case_when(asthma %in% c("NO")~ 0,
                                     asthma %in% c("YES") ~ 1),
         asthma_binary = factor(asthma_binary, 0:1, c("NO", "YES")))
           
CHIS <- CHIS %>% replace_with_na(replace = list(smoking = "INAPPLICABLE"))


table(CHIS$asthma_binary, CHIS$asthma)
```
#Checking to see see if there are any potential differences in asthma by smoking status
```{r}
CHIS %>% 
  drop_na(c(asthma_binary,smoking)) %>% 
  ggplot(aes(x = asthma_binary, y = smoking)) +
      geom_boxplot(aes(fill = asthma_binary)) +
      labs(x = "asthma Status", y = "smoking status") +
      theme_bw()
```
``
#Recoding age variable into larger intervals for logistic regression analysis
```{r}
CHIS <- CHIS %>%
  mutate(age_new = case_when(age %in% c("18-25 YEARS", "26-29 YEARS") ~ 0, age %in% c("30-34 YEARS", "35-39 YEARS") ~ 1, age %in% c("40-44 YEARS", "45-49 YEARS") ~ 2, age %in% c("50-54 YEARS", "55-59 YEARS", "60-64 YEARS", "65-69 YEARS", "70-74 YEARS", "75-79 YEARS", "80-84 YEARS", "85+ YEARS") ~ 3))

CHIS$age_new<-factor(CHIS$age_new, levels=c (0,1,2,3), labels = c("18-29", "30-39", "40-49", "50+"))

class(CHIS$age_new)
```

#Creating a new dataset with all the recoded variables and also ommiting NA values
```{r}
CHIS1<- CHIS %>%
  select("asthma_binary", "smoking", "age_new", "race") %>%
  na.omit()
```

#Table 1
```{r}
table1 (~ smoking + age_new + race | asthma_binary, data=CHIS1)
```

# 1st Logistic Model: asthma as the outcome variable and smoking status as the exposure
```{r}
smokingLogit <- glm(asthma_binary ~ smoking, data=CHIS1, family="binomial")
summary(smokingLogit)

#calculate and print ORs and 95% CIs  
ORsmoking<-exp(cbind(OR = coef(smokingLogit), confint(smokingLogit))) #calculate ORs and 95% CIs
ORsmoking #print ORs and 95% CIs

odds.n.ends (smokingLogit)

```

#Model Fits
```{r}
#influence plot - Cook's D plot-identifies observation number in parent dataset
plot(smokingLogit, which=4, id.n=5, col="red") 


#Various pseudo R squares, log likelihood, deviance, AIC, BIC
blr_model_fit_stats(smokingLogit)

#Hosmer lemeshow goodness of fit test: 
blr_test_hosmer_lemeshow(smokingLogit)

#goodness of fit test was non-significant indicating that model is a good fit
```
``
#2nd Logistic Model: with asthma as outcome, smoking as exposure, age and race as confounders.
```{r}
#income and bmi cat logistic model
smoking2Logit <- glm(asthma_binary ~ age_new + race + smoking, data=CHIS1, family="binomial")
summary(smoking2Logit)

#calculate and print ORs and 95% CIs  
ORmodel<-exp(cbind(OR = coef(smoking2Logit), confint(smoking2Logit))) #calculate ORs and 95% CIs
ORmodel #print ORs and 95% CIs

odds.n.ends(smoking2Logit)
```

Based on the data above: those who smoked some days had 2% lower odds of having asthma than those who smoked everyday (OR=0.98; 95% CI 0.73,1.32) while those who didn’t smoke at all had 10% greater odds of having asthma than those who smoked every day (OR=1.1; 95% CI 0.91,1.34). 

#Check for multicolinearity
```{r}
#Variance Inflation Factors
vif(smoking2Logit)
```

None of the values are greater than 3

#check for goodness of fit 
```{r}
#Hosmer lemeshow goodness of fit test: 
blr_test_hosmer_lemeshow(smoking2Logit)
```
Goodness of fit test was non-significant indicating that model is a good fit

#Check for undue influence
```{r}
#Cook's D plot
plot(smoking2Logit, which=4, id.n=5, col="red") 

#Cook's D cutoff=0.004
cutoff <- 0.004 

#identify observations with a Cook's Distance greater than cutoff
obs_no <- as.data.frame(cooks.distance(smoking2Logit)) %>%
  dplyr::mutate(obs_no=row_number()) %>%
  filter(`cooks.distance(smoking2Logit)` > cutoff)

#Let's exclude the values shown in the Cook's D plot, and see how the models compare.

smoking2Logit_new <- update(smoking2Logit,subset=c(-obs_no$obs_no))

summary(smoking2Logit_new)

#compare coefficients between models with and without influential observations
compareCoefs(smoking2Logit, smoking2Logit_new) 
```
After excluding values, the difference between coefficients are very little. We will continue to use the full model with all the observations for this analysis

#To compare the smoking only model with the model with smoking, age and race, we will run a LR test

```{r}
#likelihood ratio test: compare two nested models
lrtest(smoking2Logit, smokingLogit)
```
The absolute values of log likelihood of the smoking only model is larger than the model with both smoking, race and age. Based on the significant p value from LR test, we can conclude that the model with smoking, race and age is better than smoking only model.
